Score-based tests of measurement invariance: Use in practice

### 2013-10-21 frontiers/frontiers.Rnw
change according to Achim comments about results section n and sigma parameter confusion.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/semtools@278 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

To examine these specificity issues, we generated data from a
two-factor model with three indicators each (see Figure~\ref{fig:path}).
The measurement invariance
violation occurred in one of three places: the factor loading associated with Scale 1 
 ($\lambda_{11}$), the unique variance associated with Scale 1
 ($\psi_{11}$), or the factor covariance $\sigma_{12}$.  
 We then tested for measurement invariance in five subsets of parameters: the three
individual parameters noted above, all six factor loadings, % $\lambda_{11}--\lambda_{62}$, and 
and all six unique variances. % $\epsilon_{11}--\epsilon_{66}$.

@@

Full simulation results are presented in Figure~\ref{fig:sim11res} to 
Figure~\ref{fig:sim13res}; 
Figure~\ref{fig:sim11res} displays power curves as a function of 
violation magnitude for the first
factor loading $\lambda_{11}$, with panels reflecting the parameters 
being tested and lines reflecting $n$.
Figures~\ref{fig:sim12res} and Figure~\ref{fig:sim13res} display 
similar power curves when the factor covariance 
$\phi_{12}$ and error variance $\epsilon_{11}$ violate measurement
invariance, respectively. 
All three test statistics exhibit
similar results, with $\max\mathit{LM}_{uo}$ demonstrating lower power across all situations.  As described earlier, this statistic 
is asymptotically equivalent to the LRT.  $\mathit{WDM}_o$  and $\max \mathit{LM}_o$ exhibited equivalent power curves because, when only one parameter is tested, the statistics are equivalent.

@@

\begin{figure}
\caption{Simulated power curves for $\max \mathit{LM}_o$ test, $\mathit{WDM}_o$ test, and
$\mathit{LM}_{uo}$ across three levels of the ordinal variable $m$, and measurement invariance
violations of 0--4 standard errors (scaled by $\sqrt{n}$). The parameter violating measurement
invariance is $\sigma_{12}$.}
\label{fig:sim12res}
<<sim12res, fig=TRUE, height=9, width=9, echo=FALSE>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim1$test), points = TRUE, lines = TRUE)
xyplot(power ~ diff | nlevels + pars, group = ~ test, data = sim1,
       subset = (parms == "var" & 
		 pars %in% c("Covariance", "All Loadings", "All Errors") & diff %% 0.5 == 0),
       type = "b", xlab = "Violation Magnitude", ylab="Power", key = mykey, as.table = TRUE,
       strip = function(..., which.given, factor.levels){
	   if(which.given == 2){
	       strip.default(which.given, factor.levels = parlabs[c(2, 4:5)], ...)
	   } else {
	       strip.default(which.given, factor.levels = factor.levels, ...)
	   }
       })
@

@@

This Appendix includes the results for Simulation 2 when there exists model misspecification (lacking one path from Scale 1 to Math). Figure~\ref{fig:sim22res} to Figure~\ref{fig:sim24res} display results when the loading $\lambda_{11}$, covariance $\sigma_{12}$, and 
error term $\psi_{11}$ violate invariance, respectively.

@@

\begin{figure}
\caption{Simulated power curves for $\max \mathit{LM}_o$ test, $\mathit{WDM}_o$ test, and $\mathit{LM}_{uo}$ test across three levels of the ordinal variable $m$ and measurement invariance 
violations of 0--4 standard errors (scaled by $\sqrt{n}$). The violation parameter is $\sigma_{12}$.}
\label{fig:sim23res}
<<sim23res, fig=TRUE, height=9, width=9, echo=FALSE>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim2$test), points = TRUE, lines = TRUE)
xyplot(power ~ diff | nlevels + pars, group = ~ test, data = sim2,
       subset = (parms=="extra+var" & 
		 pars %in% c("Covariance", "All Loadings", "All Errors") & diff %% 0.5 == 0),
       type = "b", xlab = "Violation Magnitude", ylab = "Power", key = mykey, as.table = TRUE,
       strip = function(..., which.given, factor.levels){
	   if(which.given == 2){
	       strip.default(which.given, factor.levels = parlabs[c(2, 4:5)], ...)
	   } else {
	       strip.default(which.given, factor.levels = factor.levels, ...)
	   }
       })
@

>>>>

To examine these specificity issues, we generated data from a
two-factor model with three indicators each (see Figure~\ref{fig:path}).
The measurement invariance
violation occurred in one of three places: the factor loading associated with Scale 1 
 ($\lambda_{11}$), the unique variance associated with Scale 1
 ($\psi_{11}$), or the factor covariance $\phi_{12}$.  
 We then tested for measurement invariance in five subsets of parameters: the three
individual parameters noted above, all six factor loadings, % $\lambda_{11}--\lambda_{62}$, and 
and all six unique variances. % $\epsilon_{11}--\epsilon_{66}$.

@@

Full simulation results are presented in Figure~\ref{fig:sim11res} to 
Figure~\ref{fig:sim13res}; 
Figure~\ref{fig:sim11res} displays power curves as a function of 
violation magnitude for the first
factor loading $\lambda_{11}$, with panels reflecting the parameters 
being tested and lines reflecting different statistics.
Figures~\ref{fig:sim12res} and Figure~\ref{fig:sim13res} display 
similar power curves when the factor covariance 
$\phi_{12}$ and error variance $\epsilon_{11}$ violate measurement
invariance, respectively. 
All three test statistics exhibit
similar results, with $\max\mathit{LM}_{uo}$ demonstrating lower power across all situations.  As described earlier, this statistic 
is asymptotically equivalent to the LRT.  $\mathit{WDM}_o$  and $\max \mathit{LM}_o$ exhibited equivalent power curves because, when only one parameter is tested, the statistics are equivalent.

@@

\begin{figure}
\caption{Simulated power curves for $\max \mathit{LM}_o$ test, $\mathit{WDM}_o$ test, and
$\mathit{LM}_{uo}$ across three levels of the ordinal variable $m$, and measurement invariance
violations of 0--4 standard errors (scaled by $\sqrt{n}$). The parameter violating measurement
invariance is $\phi_{12}$.}
\label{fig:sim12res}
<<sim12res, fig=TRUE, height=9, width=9, echo=FALSE>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim1$test), points = TRUE, lines = TRUE)
xyplot(power ~ diff | nlevels + pars, group = ~ test, data = sim1,
       subset = (parms == "var" & 
		 pars %in% c("Covariance", "All Loadings", "All Errors") & diff %% 0.5 == 0),
       type = "b", xlab = "Violation Magnitude", ylab="Power", key = mykey, as.table = TRUE,
       strip = function(..., which.given, factor.levels){
	   if(which.given == 2){
	       strip.default(which.given, factor.levels = parlabs[c(2, 4:5)], ...)
	   } else {
	       strip.default(which.given, factor.levels = factor.levels, ...)
	   }
       })
@

@@

This Appendix includes the results for Simulation 2 when there exists model misspecification (lacking one path from Scale 1 to Math). Figure~\ref{fig:sim22res} to Figure~\ref{fig:sim24res} display results when the loading $\lambda_{11}$, covariance $\phi_{12}$, and 
error term $\psi_{11}$ violate invariance, respectively.

@@

\begin{figure}
\caption{Simulated power curves for $\max \mathit{LM}_o$ test, $\mathit{WDM}_o$ test, and $\mathit{LM}_{uo}$ test across three levels of the ordinal variable $m$ and measurement invariance 
violations of 0--4 standard errors (scaled by $\sqrt{n}$). The violation parameter is $\phi_{12}$.}
\label{fig:sim23res}
<<sim23res, fig=TRUE, height=9, width=9, echo=FALSE>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim2$test), points = TRUE, lines = TRUE)
xyplot(power ~ diff | nlevels + pars, group = ~ test, data = sim2,
       subset = (parms=="extra+var" & 
		 pars %in% c("Covariance", "All Loadings", "All Errors") & diff %% 0.5 == 0),
       type = "b", xlab = "Violation Magnitude", ylab = "Power", key = mykey, as.table = TRUE,
       strip = function(..., which.given, factor.levels){
	   if(which.given == 2){
	       strip.default(which.given, factor.levels = parlabs[c(2, 4:5)], ...)
	   } else {
	       strip.default(which.given, factor.levels = factor.levels, ...)
	   }
       })
@
